Supplemental Materials: Bayesian Analysis Overview

### 2023-10-23 paper/manuscript.Rmd
update abstract
minor edits

abstract: |
  Early parenting practices help shape children’s future outcomes. In particular, high-quality early interactions can facilitate language learning and school performance. Phone-based parenting applications can deliver low-cost interventions on parenting style, yet little is known about the efficacy of this approach. In two studies (N=142; 79 female; ages 6-24 months; 74 white, 31 Asian, 31 biracial, 3 black, 12 other; 31 Hispanic), we investigated effects of short videos on parent behavior. After parents watched videos depicting age-appropriate parent-child activities, parents spoke more (d=0.83) and made more bids for joint attention (d=1.03). These results suggest that activity videos can lead to local changes in parent engagement, providing support for the use of such videos as a part of broader app-based parenting interventions.
keywords          : "language development; parent intervention; childhood development; joint attention; lexical diversity"
wordcount         : "5509"
bibliography      : ["library.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
header-includes:
  - \usepackage{float}
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

@@

The quantity and quality of early language input has been found to be strongly associated with later language and academic outcomes [@Hart1995; @Marchman2008; @Rowe2012; @Cartmill2013; @HirshPasek2015; see @Schwab2016 for a review]. 
Thus, because of the potential for large downstream effects [@Heckman2006], there is tremendous interest in interventions that change children's language environment. 
And because parents define a large portion of that environment, especially before the onset of formal schooling, parent behavior is a critical locus for such interventions. 
Many effective parenting interventions require large resource investments and require many hours of in-person contact [@Jamaica2014; @PerryPreschool2004; @Leung2018], making implementation at scale a daunting proposition. 
For this reason, many researchers targeting early language are interested in delivering parenting interventions remotely -- through texts, apps, and videos delivered on digital devices. 
But what do parents take away from these short messages about what to do with or how to talk with their children? 
In this paper, rather than conducting a large-scale intervention, we investigate the ways that parents change their local behavior with children immediately after seeing a short video instructing them on an activity, investigating one of the building blocks of many larger interventions.

@@

But is this theory of change correct in the case of digital parenting interventions? 
That is, does the provision of knowledge and activities via short videos or messages actually lead to higher-quality play? 
Alternatively, this approach could be flawed. 
For example, focusing parents on a specific activity could cause them parents to over-focus on achieving the superficial goals of the activity while neglecting the broader goals of fostering more -- and more responsive -- interaction. 
This problem might be especially likely with video messages, which could encourage parents to try to mimic a model's specific speech and/or actions.
Attempting to reproduce the surface details of a video-guided activity could in turn result in less high-quality talk, with less responsiveness to their child's play, a target of parent responsiveness interventions [e.g., @Landry2008].
(Speaking as a parent, one author of the current paper has found himself occasionally so engrossed in attempting to execute an activity with his children that he has lost track of whether the children are continuing to enjoy it!).

Thus, our goal was to gather information about whether the use of short video parenting messages as a component of broader interventions is supported by their local effects on parent-child interaction and language use. 
Our current studies were designed to assess how parents change their interactions with young children on the basis of these videos.  
We made a first step towards assessing this question by performing two studies within the context of a convenience sample of parent-child dyads in a local children's museum. 
We showed parents in the experimental group a single short video modeling an interactive toy-based activity along with a scientific justification. 
Parents in the control group received either no video (Experiment 1) or a video of a recent finding in developmental psychology (Experiment 2). 
We then gave the toys from the video to all dyads and videotaped their interactions, coding for caregivers' language quantity and quality as well as joint attention.

@@

To get a better sense of the intervention's effect on language use, we analyzed which words were characteristic of parents' speech in each condition, comparing the difference in frequency rank of each word (lemma) in the two conditions, as well as contrasting the corpus overall with a general English-language word frequency list ([see SI](https://osf.io/2bpdf/files/github/language_analyses/Exp1_characteristic_chart.html) for the interactive corpus characteristic plot). 
Words that were strongly indicative of being from the Activity Video condition include "give", "big", "small", "ribbit", "thank", "have", and "bus", while words that were most characteristic of the No Video condition include "shake", "ready", "oh", "on", "going", "going", "see", "let", and the child's name.

>>>>

abstract: |
  Early parenting practices help shape children’s future outcomes. In particular, high-quality early interactions can facilitate language learning and school performance. Phone-based parenting applications can deliver low-cost interventions on parenting style. In two studies (N=142; 79 female; ages 6-24 months; 74 white, 31 Asian, 31 biracial, 3 black, 12 other; 31 Hispanic), we investigated effects of short videos on parent behavior. Immediately after parents watched videos depicting age-appropriate parent-child activities, parents spoke more (d=0.83) and made more bids for joint attention (d=1.03). These results suggest that activity videos can lead to near-term changes in parent engagement, providing support for the use of such videos as a part of broader app-based parenting interventions.
keywords          : "language development; parent intervention; childhood development; joint attention; lexical diversity"
wordcount         : "5509"
bibliography      : ["library.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
header-includes:
  - \usepackage{float}
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

@@

The quantity and quality of early language input has been found to be strongly associated with later language and academic outcomes [@Hart1995; @Marchman2008; @Rowe2012; @Cartmill2013; @HirshPasek2015; see @Schwab2016 for a review]. 
Thus, because of the potential for large downstream effects [@Heckman2006], there is tremendous interest in interventions that change children's language environment. 
And because parents define a large portion of that environment, especially before the onset of formal schooling, parent behavior is a critical locus for such interventions. 
Many effective parenting interventions require large resource investments and require many hours of in-person contact [@Jamaica2014; @PerryPreschool2004; @Leung2018], making implementation at scale a daunting proposition. 
For this reason, many researchers targeting early language are interested in delivering parenting interventions remotely -- through texts, apps, and videos delivered on digital devices. 
But what do parents take away from these short messages about what to do with or how to talk with their children? 
In this paper, rather than conducting a large-scale intervention, we investigate the ways that parents change their local (i.e. near-term) behavior with children immediately after seeing a short video instructing them on an activity, investigating one of the building blocks of many larger interventions.

@@

But is this theory of change correct in the case of digital parenting interventions? 
That is, does the provision of knowledge and activities via short videos or messages actually lead to higher-quality play? 
Alternatively, this approach could be flawed. 
For example, focusing parents on a specific activity could cause parents to over-focus on achieving the superficial goals of the activity while neglecting the broader goals of fostering more -- and more responsive -- interaction. 
This problem might be especially likely with video messages, which could encourage parents to try to mimic a model's specific speech and/or actions.
Attempting to reproduce the surface details of a video-guided activity could in turn result in less high-quality talk, with less responsiveness to their child's play, a target of parent responsiveness interventions [e.g., @Landry2008].
(Speaking as a parent, one author of the current paper has found himself occasionally so engrossed in attempting to execute an activity with his children that he has lost track of whether the children are continuing to enjoy it!).

Thus, our goal was to gather information about whether the use of short video parenting messages as a component of broader interventions is supported by their local effects on parent-child interaction and language use. 
Our current studies were designed to assess how parents change their near-term interactions with young children on the basis of these videos. 
We made a first step towards assessing this question by performing two studies within the context of a convenience sample of parent-child dyads in a local children's museum. 
We showed parents in the experimental group a single short video modeling an interactive toy-based activity along with a scientific justification. 
Parents in the control group received either no video (Experiment 1) or a video of a recent finding in developmental psychology (Experiment 2). 
We then gave the toys from the video to all dyads and videotaped their interactions, coding for caregivers' language quantity and quality as well as joint attention.

@@

To get a better sense of the intervention's effect on language use, we analyzed which words were characteristic of parents' speech in each condition, comparing the difference in frequency rank of each word (lemma) in the two conditions, as well as contrasting the corpus overall with a general English-language word frequency list ([see SI](https://osf.io/2bpdf/files/github/language_analyses/Exp1_characteristic_chart.html) for the interactive corpus characteristic plot). 
Words that were strongly indicative of being from the Activity Video condition include "give", "big", "small", "ribbit", "thank", "have", and "bus", while words that were most characteristic of the No Video condition include "shake", "ready", "oh", "on", "going", "see", "let", and the child's name.
